---
title: "**W271**-2 -- Spring 2016 -- **HW 8**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*April 6, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW 8 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
---

**********

\pagebreak

# Exercises

```{r, echo = FALSE, warning = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW8_template.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```

## Exercise 1

**Build an univariate linear time series model (i.e AR, MA, and ARMA models) using the series in `hw08_series.csv`.**

+ **Use all the techniques that have been taught so far to build the model, including date examination, data visualization, etc.**

+ **All the steps to support your final model need to be shown clearly.**

+ **Show that the assumptions underlying the model are valid.**

+ **Which model seems most reasonable in terms of satisfying the model’s underling assumption?**

+ **Evaluate the model performance (both in- and out-of-sample)**

+ **Pick your "best" models and conduct a 12-step ahead forecast. Discuss your results. Discuss the choice of your metrics to measure "best".**

**********
