---
title: "**W271**-2 -- Spring 2016 -- **HW 2**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*February 10, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW2 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}


---

**********

# Data

**In the United States, a 401K is a type of retirement savings plan that is tied to a worker's place of employment. Employees that put money into a 401K enjoy certain tax benefits. Moreover, many employers have a policy of promoting 401K use, by matching some percentage of an employee's contributions. If an employer matches at, say, 50%, for every dollar that an employee puts into a 401K, the employer will put in another 50 cents.**

**The file `401k_w271.RData` contains data on 401k contributions that were filed with the IRS on form 5500. It was collected by Professor L. E. Papke and may have been further modified by the instructors to test your proficiency.**


\pagebreak

# Exercises

**Complete the following exercises, following the best practices outlined in class. Place your answers in a written report (pdf, word, or jupyter notebook format) along with relevant R statements and output.**

**Load the `401k_w271.RData` dataset and look at the value of the function `desc()` to see what variables are included.**

```{r, echo = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW2_AL.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
#opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```

```{r Load_Data, echo = -c(1:2)}
```



## Question 1

**Your dependent variable will be `prate`, representing the fraction of a company's employees participating in its 401k plan. Because this variable is bounded between 0 and 1, a linear model without any transformations may not be the most ideal way to analyze the data, but we can still learn a lot from it. Examine the `prate` variable and comment on the shape of its distribution.**

```{r Question1-1, echo = -c(1:2)}
```

Based on the plots above, the data is very negativley skewed with a large number of observations with close to 100% participation. Additionally, from the initial data examination, there were no NA values and there are three data points that are greater than 100, which does not make sense in the context of the data. The second plot excludes these data points and the analysis in subsequent questions will also excludes these outliers.  


## Question 2

**Your independent variable will be `mrate`, the rate at which a company matches employee 401K contributions. Examine this variable and comment on the shape of its distribution.**

```{r Question2-1, echo = -c(1:2)}
```

Based on the historgram above the mrate data is positively skewed with a right tail. A large portion of the observations are near zero. Similarly to the prate variable there are no NA values. We are assuming the zero values are reported zeros and not NA substitution. 


\pagebreak

## Question 3

**Generate a scatterplot of `prate` against `mrate`. Then estimate the linear regression of `prate` on `mrate`. What slope coefficient did you get?**

```{r Question3-1, echo = FALSE, fig.width = 6, fig.height = 4.5, fig.cap = "Scatterplot of the participation rate ($\\%$) to 401K plans of a company's employees against the match rate ($\\%$) of that company to their employees' contributions"}
```

The scatterplot shows that the higher the match rate by a company corresponds to a higher particpation rate, whereas a low match rate has a wide spread of particpation rates. 

```{r Question3-2, echo = c(1:2)}
```
The linear regression has a slope coefficient os 5.86 indicating a positve relationship with increasing match rate will increase participation. 



## Question 4

**Is the assumption of zero-conditional mean realistic? Explain your evidence. What are the implications for your OLS coefficients?**

We can examine the zero-conditional mean assumption with a residual vs fitted plot. 

```{r Question4, echo = -c(1:2)}
```


## Question 5

**Is the assumption of homoskedasticity realistic? Provide at least two pieces of evidence to support your conclusion. What are the implications for your OLS analysis?**

```{r Question5, echo = -c(1:2)}
```

The residuals vs fitted plot can also provide insights on the homoskedasticity assumption. In this case this assumption is not realisitic. As seen in the residuals vs fitted plot, the data is very widely dispersed on the left side and funnels towards the right. This indicates that the variation accross all values is not even which indicates we have heteroskedasticity. 

The implicaitons of heterskedasticity are the coefficients from the OLS may be pulled with the larger variations and the standard error estimate will be biased. The use of robust standard error calcualations or weighted least squares regression can help correct this bias, but may necessitate other assumptions to be met.  Alternatively, a transform (ie log or square root) could be applied to the variable to try and address the heteroskedasticity. However, with the large sample size, failure to meet the homoskedasticity assumption does not impact the OLS results dramatically. 


## Question 6

**Is the assumption of normal errors realistic? Provide at least two pieces of evidence to support your conclusion. What are the implications for your OLS analysis?**



## Question 7

**Based on the above considerations, what is the standard error of your slope coefficient?**

Based on the above considerations, the standard error of the slope coefficient may be biased because the assumptions are not met. However, we can get the standard error of the slope coeffient form the output of the linear model. 

```{r Question7, echo = -c(1:2)}

```


## Question 8

**Is the effect you find statistically significant, and is it practically significant?**

```{r Question8, echo = -c(1:2)}
```

The linear model shows a p-value <.05 which indicates a statistically significant result. This implies would could reject the null hypothesis. However, the size of the dataset can cause the restult to be statistically significant without having any practical significance. In this case, the result may be practically significant as well indicating a positive relationship between the amount a company matches increasing the amount of participation in the 401k plans. 
