---
title: '**W271**-2 -- Spring 2016 -- **HW 2**'
author: '***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***'
date: '*February 10, 2016*'
output:
  pdf_document:
    fig_caption: yes
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW2 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
linkcolor: blue
numbersections: no
options: width=30
geometry: margin=1in
fontsize: 10pt
---

**********


# Exercises

**Complete the following exercises, following the best practices outlined in class. Place your answers in a written report (pdf, word, or jupyter notebook format) along with relevant R statements and output.**

```{r, echo = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW3_AL.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```



## Question 1

**Load the twoyear.RData dataset and describe the basic structure of the data.**


```{r Load_Data, echo = -c(1:2)}
```

```{r Question1-1, echo = -c(1:2)}
```



## Question 2

**Typically, you will need to thoroughly analyze each of the variables in the data set using univariate, bivariate, and multivariate analyses before attempting any model. For this homework assume that this step has been conducted. Estimate the following regression:**

$$\begin{aligned}\mathbf{log(wage) =}& \mathbf{\ \beta_0 + \beta_1jc + \beta_2univ + \beta_3exper + \beta_4black + \beta_5hispanic} \\ & \mathbf{+\ \beta_6AA + \beta_7BA + \beta_8exper \cdot black+e}\end{aligned}$$
    
**Interpret the coefficients $\hat{\beta}_4$ and $\hat{\beta}_8$.**


```{r Question2, echo = -c(1:2)}
```




\pagebreak

## Question 3

**With this model, test that the return to university education is 7%**


```{r Question3, echo = -c(1:2)}
```

The coefficient for the univ variable is .073 which equates to approximately a 7% increase in log(wage) for every increment increase in univ education. 

## Question 4

**With this model, test that the return to junior college education is equal for black and non-black.**

```{r Question4, echo = -c(1:2)}
```


## Question 5

**With this model, test whether the return to univeristy education is equal to the return to 1 year of working experience.**


```{r Question5, echo = -c(1:2)}
```


## Question 6

**Test the overall significance of this regression.**

```{r Question6, echo = -c(1:2)}
```

The overall significance of this regression is high with a F-statistic of 249.6 and an overall p-value <2.2e-16 indicating that overall this model is performing better than random. However, we can likely increase our r squared value by eliminating some of the factors  because several factors individually are non significant. 

## Question 7

**including a square term of working experience to the regression model built above, estimate the linear regression model again. What is the estimated return to work experience in this model?**

```{r Question7, echo = -c(1:2)}
```


## Question 8

**Provide the diagnosis of the homoskedasticity assumption. Does this assumption hold? If so, how does it affect the testing of no effect of university education on salary change? If not, what potential remedies are available?**

```{r Question8, echo = -c(1:2)}
```


The homoskedasticity assumption may hold in this case. Although the residuals vs fitted plot form the model appears to have a little less variance on the edges (far right and far left) the red line is relatively straight and flat. 