---
title: '**W271**-2 -- Spring 2016 -- **HW 2**'
author: '***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***'
date: '*February 10, 2016*'
output:
  pdf_document:
    fig_caption: yes
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW2 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
linkcolor: blue
numbersections: no
options: width=30
geometry: margin=1in
fontsize: 10pt
---

**********


# Exercises

**Complete the following exercises, following the best practices outlined in class. Place your answers in a written report (pdf, word, or jupyter notebook format) along with relevant R statements and output.**

```{r, echo = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW2_Carin.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```



## Question 1

**Load the twoyear.RData dataset and describe the basic structure of the data.**


```{r Load_Data, echo = -c(1:2)}
```

```{r Question1-1, echo = -c(1:2)}
```



## Question 2

**Typically, you will need to thoroughly analyze each of the variables in the data set using univariate, bivariate, and multivariate analyses before attempting any model. For this homework assume that this step has been conducted. Estimate the following regression:**

$$\log(wage)=\Beta_0+\Beta_1 jc+\Beta_2 univ+\Beta_3 exper+\Beta_4 black+\Beta_5 hispanic+\Beta_6 AA+\Beta_7 BA+\Beta_8 exper\cdot black+\epsilon$$
  
**Interpret the coefficients $\hat{\Beta_4}$ and $\hat{\Beta_8}$.**


```{r Question2-1, echo = -c(1:2)}
```


\pagebreak

## Question 3

**With this model, test that the return to university education is 7%**


```{r Question3, echo = -c(1:2)}
```


## Question 4

**With this model, test that the return to junior college education is equal for black and non-black.**

```{r Question4, echo = -c(1:2)}
```


## Question 5

**With this model, test whether the return to univeristy education is equal to the return to 1 year of working experience.**


```{r Question5, echo = -c(1:2)}
```


## Question 6

**Test the overall significance of this regression.**

```{r Question6, echo = -c(1:2)}
```

## Question 7

**including a square term of working experience to the regression model built above, estimate the linear regression model again. What is the estimated return to work experience in this model?**

```{r Question7, echo = -c(1:2)}
```


## Question 8

**Provide the diagnosis of the homoskedasticity assumption. Does this assumption hold? If so, ow does it affect the testing of no effect of university education on salary change? If not, what potential remedies are available?**

```{r Question8, echo = -c(1:2)}
```
