---
title: "**W271**-2 -- Spring 2016 -- **HW 3**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*February 17, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW2 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
---
  
**********



\pagebreak

# Exercises

```{r, echo = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW3_Carin.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```



## Question 1

**Load the `twoyear.RData` dataset and describe the basic structure of the data.**
  
```{r Question1, echo = -c(1:4)}
```



**********

## Question 2

**Typically, you will need to thoroughly analyze each of the variables in the data
set using univariate, bivariate, and multivariate analyses before attempting any
model. For this homework, assume that this step has been conducted. Estimate
the following regression:**
  
$$\begin{aligned}\mathbf{log(wage) =}& \mathbf{\ \beta_0 + \beta_1jc + \beta_2univ + \beta_3exper + \beta_4black + \beta_5hispanic} \\ & \mathbf{+\ \beta_6AA + \beta_7BA + \beta_8exper \cdot black+e}\end{aligned}$$
  
**Interpret the coefficients $\hat{\beta}_4$ and $\hat{\beta}_8$.**
  
```{r Question2, echo = -c(1:2)}
```



**********

## Question 3

**With this model, test that the return to university education is $7\%$.**
  
```{r Question3, echo = -c(1:2)}
```



**********

## Question 4

**With this model, test that the return to junior college education is equal for black and non-black.**
  
```{r Question4, echo = -c(1:3)}
```



**********

## Question 5

**With this model, test whether the return to university education is equal to the return to 1 year of working experience.**
  
```{r Question5, echo = -c(1:3)}
```



**********

## Question 6

**Test the overall significance of this regression.**
  
```{r Question6, echo = -c(1:2)}
```



**********

## Question 7

**Including a square term of working experience to the regression model built above, estimate the linear regression model again. What is the estimated return to work experience in this model?**
  
```{r Question7, echo = -c(1:4)}
```



**********

\pagebreak

## Question 8

**Provide the diagnosis of the homoskedasticity assumption. Does this assumption hold? If so, how does it affect the testing of no effect of university education on salary change? If not, what potential remedies are available?**

```{r Question8, echo = FALSE, results='asis'}
```

...

```{r Question8-2, echo = FALSE, results='asis'}
```

...

**********
