---
title: "**W271**-2 -- Spring 2016 -- **HW 7**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*March 30, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW 7 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
---

**********

\pagebreak

# Exercises

```{r, echo = FALSE, warning = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW7_Carin.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```

## Question 1

1.1. **Load `hw07_series1.csv`.**

```{r Question1-1, echo = -c(1:3)}
```


**********

1.2. **Describe the basic structure of the data and provide summary statistics of the series.**

```{r Question1-2, echo = -c(1:2)}
```

The data correspond to `r dim(hw07)[1]` observations of a single variable. No information about the time scale is given, so we'll just use an index between 1 and `r dim(hw07)[1]` (with `frequency = 1`). The main descriptive statistics are shonw in the Table above.


**********

\pagebreak

1.3. **Plot histogram and time-series plot of the series. Describe the patterns exhibited in histogram and time-series plot. For time series analysis, is it sufficient to use only histogram to describe a series?**

```{r Question1-3-1, echo = 3, fig.width=6, fig.height=4.5, fig.cap = "Histogram of the data"}
```

Let's get some descriptive statistics about the distribution of the data, that complement the histogram above:

```{r Question1-3-2, echo = FALSE}
```

The series has an *excess kurtosis* (kurtosis minus 3, the value for a normal distribution) which is negative, which indicates a platykurtic distribution (thinner tails than a normal one). But that excess kurtosis is close to zero, and not statistically significant (the parameter `kurt.2SE` would have to be greater than 1; see the [documentation of `stat.desc`](http://127.0.0.1:13524/library/pastecs/html/stat.desc.html) for more information). Likewise, the *skewness* is positive, which indicates a right-skewed distribution (with a heavy right tail), but again is not far from zero and not statistically significantly different than it. A Shapiro-Wilk test is also non-significant ($p = `r more_desc_stat[which(rownames(more_desc_stat) == 'normtest.p')]`$), so we cannot reject the hypothesis that the distribution of the series is normal (though its size---`r dim(hw07)[1]`---is a bit reduced).

For time series analysis (unlike cross-sectional data analysis), the histogram alone is not enough to describe a series because it tells us nothing about the dynamics of it; e.g., this one in particular lets us know that there was one time period where the value of the time series was below 10, but not when that happened (in the 4th time period, as shown in the next Figure).

```{r Question1-3-3, echo = FALSE, fig.width=6, fig.height=4.5, fig.cap = "Time-series plot of the data"}
```

The time-series plot has been plotted together with the mean value and a 9-point symmetric moving average, that exhibits an increasing trend during the first 30 time periods or so, then a small decline for about 10 time periods, another increasing trend for a few time periods, and finally a decreasing trend during the last 30 time periods or so (this one not so linear; the decrease is steeper at the beginning, mainly due to the effect of a pike in observations 44 and 45); the smoother yields no values for the first and last 4 observations, but it seems the trend becomes positive again at the end of the observation period. Overall, we can say the time series is persistent.


**********

\pagebreak

1.4. **Plot the ACF and PACF of the series. Describe the patterns exhibited in the ACF and PACF.**

```{r Question1-4, echo = FALSE, fig.width=6, fig.height=4.5, fig.cap = "Autocorrelation and partial autocorrelation graphs"}
```

The correlogram has a wave-like shape that resembles that of a shrinking cosine function (typical of an AR(2) process). It decreases relatively slowly (the first autocorrelation not statistically significantly different from zero corresponds to the 9th lag). As for the PACF, it drops off relatively abruptly at lag 2 (another indication of an AR(2) process).


**********

1.5. **Estimate the series using the `ar()` function.**

```{r Question1-5, echo = c(2:4, 6:9)}
```


**********

1.6. **Report the estimated AR parameters, the order of the model, and standard errors.**

Order of the model:

```{r Question1-6-1, echo = -1}
```

Estimated AR parameters:

```{r Question1-6-2}
```

Other parameters of the estimated AR model:

```{r Question1-6-3}
```

As shown in the output of the code above, models of other orders (especially the one of order 3) have similar AIC values. The last parameter shown above is the asymptotic covariance matrix of the coefficient estimates, so the square root of the elements in the diagonal of that matrix are the standard errors. Together with them, we can also estimate the confidence interval of the coefficient estimates:

```{r Question1-6-4, echo = 1}
```

Both coefficients are significant (the CI does not include zero in both cases).

**********
**********

\pagebreak



## Question 2

2.1. **Simulate a time series of lenght 100 for the following model. Name the series $x$.**

$$\mathbf{x_t = \frac{5}{6} x_{t-1} - \frac{1}{6} x_{t-2} + \omega_t}$$

This is an AR(2) model with coefficients $5/6 = `r frmt(5/6)`$ and $1/6 = `r frmt(1/6)`$ respectively.

```{r Question2-1, echo = c(4:5), fig.width = 6, fig.height = 6, fig.cap = "Histogram and time-series plot of the simulated time series"}
```

> (The Figure above have the same legends than Figures 1 and 2, Question 1.)


**********

\pagebreak

2.2. **Plot the correlogram and partial correlogram for the simulated series. Comments on the plots.**

```{r Question2-2, echo = FALSE, fig.width=6, fig.height=4.5, fig.cap = "Autocorrelation and partial autocorrelation graphs"}
```

As in Question 1, the correlogram has a wave-like shape that resembles that of a shrinking cosine function (though the sign remains negative for the 3rd semi-period). IT and it decreases relatively quickly (though this changes in other simulations---if other seed is used).

The PACF drops off relatively abruptly at lag 1 (or lag 2, in---a few of the--other simulations).


**********

\pagebreak

2.3. **Estimate an AR model for this simulated series. Report the estimated AR parameters, standard errors, and the order of the AR model.**

```{r Question2-3}
```


**********

2.4. **Construct a 95% confidence intervals for the parameter estimates of the estimated model. Do the "true" model parameters fall within the confidence intervals? Explain the 95% confidence intervals in this context.**

See Table 4. Yes, they do. What happens in many of the simulations (when using a different seed is that the best fitted model is an AR(1) so there is no estimate for the coefficient for lag 2). In any case, if we repeat the simulation over and over, the "true" model parameters will fall within the CIs about 95% of the times (at least the coefficient of lag 1; and also the coefficient of lag 2 if we "force" our fitted model to be an AR(2)).

...

**********

2.5. **Is the estimated model stationary or non-stationary?**

...

**********

2.6. **Plot the correlogram of the residuals of the estimated model. Comment on the plot.**

...

**********
**********
