---
title: "**W271**-2 -- Spring 2016 -- **HW 7**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*March 30, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW 7 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
---

**********

\pagebreak

# Exercises

```{r, echo = FALSE, warning = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW7_Carin.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```

## Question 1

1.1. **Load `hw07_series1.csv`.**

```{r Question1-1, echo = -c(1:3)}
```


1.2. **Describe the basic structure of the data and provide summary statistics of the series.**

...

1.3. **Plot histogram and time-series plot of the series. Describe the patterns exhibited in histograrm and time-series plot. For time series analysis, is it sufficient to use only histogram to describe a series?**

...

1.4. **Plot the ACF and PACF of the series. Describe the patterns exhibited in the ACF and PACF.**

...

1.5. **Estimate the series using the `ar()` function.**

...

1.6. **Report the estimated AR parameters, the order of the model, and standard errors.**

...

**********

\pagebreak



## Question 2

2.1. **Simulate a time series of lenght 100 for the following model. Name the series $x$.**

$$\mathbf{x_t = \frac{5}{6} x_{t-1} - \frac{1}{6} x_{t-2} + \omega_t}$$

...

2.2. **Plot the correlogram and partial correlogram for the simulated series. Comments on the plots.**

...

2.3. **Estimate an AR model for this simulated series. Report the estimated AR parameters, standard errors, and the order of the AR model.**

...

2.4. **Construct a 95% confidence intervals for the parameter estimates of the estimated model. Do the "true" model parameters fall within the confidence intervals? Explain the 95% confidence intervals in this context.**

...

2.5. **Is the estimated model stationary or non-stationary?**

...

2.6. **Plot the correlogram of the residuals of the estimated model. Comment on the plot.**

...

**********
