---
title: "**W271**-2 -- Spring 2016 -- **Lab 2**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*March 7, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- Lab 2}
- \fancyhead[RE,RO]{\leftmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
---

**********

\pagebreak

```{r, echo = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_Lab2_Carin - Q6 only.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```

# Question 6: CLM 3

**Your analytics team has been tasked with analyzing aggregate revenue, cost and sales data, which have been provided to you in the R workspace/data frame `retailSales.Rdata`.**

**Your task is two fold. First, your team is to develop a model for predicting (forecasting) revenues. Part of the model development documentation is a backtesting exercise where you train your model using data from the first two years and evaluate the model's forecasts using the last two years of data.**

**Second, management is equally interested in understanding variables that might affect revenues in support of management adjustments to operations and revenue forecasts. You are also to identify factors that affect revenues, and discuss how useful management's planned revenue is for forecasting revenues.**

**Your analysis should address the following:**

+ **Exploratory Data Analysis: focus on bivariate and multivariate relationships.**

+ **Be sure to assess conditions and identify unusual observations.**

First we explore the whole dataset.

> ```{r Question6-1, echo = -c(1:2)}
> ```

The dataset contains `r frmt(dim(data)[1], 0)` observations of `r dim(data)[2]` variables. `r sum(sapply(data, is.factor))` of them are categorical (``r paste(names(data)[sapply(data, is.factor)], collapse=", ")``), and `Year` should also be considered as categorical, since there are data from only `r max(data$Year) - min(data$Year) + 1` years (from `r min(data$Year)` to `r max(data$Year)`).

> ```{r Question6-2}
> ```

We also notice (from the output of `summary`) that some of the variables (all of them numerical) has a high number of `NA`s, the same in all cases (`r sum(is.na(data$Revenue))`, i.e., $`r frmt(sum(is.na(data$Revenue))/dim(data)[1]*100, 2)`\%$ of the total number of observations). Do the `NA`s appear in the same observations for all those variables? Yes, they do.

> ```{r Question6-3}
> ```

And the amount of `NA`s per category is roughly the same for all categorical values (or at least there are non-missing data for all categories; below we just show the percentage per category for three of the numerical variables).

> ```{r Question6-4, echo = c(1:3)}
> ```

So we can ommit all those missing observations (reducing our sample size to `r dim(data)[1] - sum(is.na(data$Revenue))`), and continue with a further analysis of the numerical variables:

> ```{r Question6-5}
> ```

All numerical variables are right-skewed, with long right tails (i.e., with several observations more than 2 standard deviations far from the mean), especially the ones corresponding to aggregate---non-unit---results.

> ```{r Question6-6, echo = FALSE, fig.width = 6, fig.height = 4.5, fig.cap = "Histogram of all non-categorical variables in the dataset"}
> ```

Below we show the correlation matrix of the numerical variables, as well as two different representations of the scatterplot matrix (where we've used a sample of the data of size 500 because the plotting functions consume a lot of resources; that's why the correlations shown in the second Figure, only approximate, differ from the ones shown right below). As we might have expected, the correlations between `Revenue`, `Planned.revenue`, `Product.cost`, and `Gross.profit` (i.e., the aggregate values), as well as those between `Unit.cost`, `Unit.price`, and `Unit.sale.price` (i.e., the values per unit), are positive and very high. `Quantity` is negatively correlated with the unitary variables (but that correlation is negligible in absolute value), and is moderately correlated ($\rho \simeq 0.5$) with the aggregate values.

> ```{r Question6-7, echo = 1, fig.width = 6, fig.height = 6, fig.cap = "Scatterplot matrix of a sample of the dataset"}
> ```

> ```{r Question6-8, echo = FALSE, fig.width = 6, fig.height = 6, fig.cap = "Scatterplot matrix of a sample of the dataset (with correlations)"}
> ```

\pagebreak

  After our EDA, we can divide the dataset into two separate ones, to train and evaluate the model. Now we can convert `Year` back to a numerical variable (subtracting 2004 so the baseline is 0; that will make the intercept more intuitive when including `Year` in the regression model).

> ```{r Question6-9}
> ```

Not all products appear in both periods so some re-factoring is needed:

> ```{r Question6-10}
> ```

There are some variables that are calculated from `Revenue` (or vice versa) so including them in any regression model would lead to a perfect fit. In particular, `Gross.profit` = `Revenue` - `Product.cost`. And `Revenue` should be equal to `Unit.sale.price` times `Quantity`, though this is not always the case, and there are differences in many cases ($`r frmt(100*sum(data$Unit.sale.price * data$Quantity != data$Revenue) / dim(data)[1], 1)` \%$ of the total number of observations).

> ```{r Question6-11}
> ```

So `Revenue` and `Product.cost` should definitely not be included in the regression model, but `Unit.sale.price` and `Quantity` might.

Let's start with the simplest model:

> ```{r Question6-12, echo = c(1:5), fig.cap = "Planned Revenue vs. Revenue (observed and predicted) in 2006 and 2007"}
> ```

We'll use the RMSE to compare different models:

> ```{r Question6-13}
> ```

+ **Is the change in the average revenue different from 95 cents when the planned revenue increases by $1?**

As shown below, the change in the average revenue is significantly different from $0.95 when the revenue increases by $1 (while the *F* statistic of the exact value, which is quite close to $0.95, has a *p* value equal to 1):

> ```{r Question6-14}
> ```

> ```{r Question6-15, echo = c(1:5), fig.cap = "Planned Revenue vs. Revenue (observed and predicted) in 2006 and 2007"}
> ```

+ **Explain what interaction terms in your model mean in context supported by data visualizations.**

+ **Give two reasons why the OLS model coefficients may be biased and/or not consistent, be specific.**

+ **Propose (but do not actually implement) a plan for an IV approach to improve your forecasting model.**


**********
