---
title: "**W271**-2 -- Spring 2016 -- **HW 5**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*March 9, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW 5 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
---

**********

\pagebreak

# Exercises

```{r, echo = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW5_Carin.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```



## Question 1

1. **Install the library `"astsa"` using the function: `install.packages("astsa")`**

```{r Question1-1, echo = -c(1:2)}
```

2. **Load the library: `library(astsa)`**

```{r Question1-2, echo = TRUE}
```

3. **Use the function `str()` to see the information of a particular data series, such as `str(EQ5)` for the Seismic Trace of Earthquake number 5 series**

```{r Question1-3, echo = -c(1:2)}
```

According to that [package documentation](https://cran.r-project.org/web/packages/astsa/astsa.pdf), `EQ5` corresponds to the *Seismic trace of an earthquake [two phases or arrivals along the surface, the primary wave (t = 1, . . . , 1024) and the shear wave (t = 1025, . . . , 2048)] recorded at a seismic station.*

4. **Plot the time series plots and histograms of the following 3 series. Feel free to use the codes provided in the R scripts. Make sure that each of your graph has a title, the axis ticks are clear, the axes are well-labelled, and use color intelligently.**

```{r Question1-3, echo = FALSE}
```

5. **Write a few sentences to describe each of the series.**

  + **EQ5**

  ...

  + **flu**

  ...

  + **gas**

  ...



**********

## Question 2

**Describe 3 examples you have used in your work or encounter in real life. Ideally, you can even load at least one of these time series, plot it, and the write a few statements to describe its characteristics.**


```{r Question2, echo = -c(1:4)}
```



**********

## Question 3

**Simulate a white noise series with 1000 random draws and plot (1) a time series plot and (2) a histogram. The usual reqirements on graphics (described) in Question 1) applied.**

```{r Question3, echo = -c(1:4)}
```



**********

## Question 4

**Simulate (with 1000 random draws) the following two zero-mean autoregressive model with order 1 (i.e. AR(1)) models:**

$$\mathbf{y_t = 0.9y_{t1} + w}$$

$$\mathbf{y_t = 0.2y_{t1} + w}$$

**Plot a time plot for each of the simulated series. Graph a histogram for each of the simulated series. Write a few statements to compare the two series.**

```{r Question4, echo = -c(1:7)}
```



**********

## Question 5

**Simulate (with 1000 random draws) the following 3 models:**

1. **A deterministic linear (time) trend of the form: $y_t = 10 + 0.5t$**

2. **Random walk without drift**

3. **Random walk with drift = 0.5**

**Plot a time plot for each of the simulated series. Graph a histogram for each of the simulated series. Write a few statements to compare the two series.**

```{r Question5, echo = -c(1:7)}
```

**********
