---
title: "**W271**-2 -- Spring 2016 -- **HW 4**"
author: "***Juanjo Carin, Kevin Davis, Ashley Levato, Minghu Song***"
date: "*February 24, 2016*"
output:
   pdf_document:
     fig_caption: yes
     toc: yes
numbersections: false
geometry: margin=1in
options: width=30
fontsize: 10pt
linkcolor: blue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[LO,LE]{Carin, Davis, Levato, Song}
- \fancyhead[CE,CO]{W271 -- HW2 -- \leftmark}
- \fancyhead[RE,RO]{\rightmark}
- \fancyfoot[LO,LE]{UC Berkeley -- MIDS}
- \fancyfoot[CO,CE]{Spring semester 2016}
- \fancyfoot[RE,RO]{\thepage}
- \renewcommand{\headrulewidth}{0.5pt}
- \renewcommand{\footrulewidth}{0.5pt}
---


\pagebreak

# Exercises

```{r, echo = FALSE}
require(knitr, quietly = TRUE)
read_chunk('code/W271_HW5_template.R')
opts_chunk$set(message = FALSE, warning = FALSE)
opts_chunk$set(fig.width = 4, fig.height = 3)
# Set path to data here (don't use setwd() inside a chunk!!!)
opts_knit$set(root.dir = './data')
```

```{r Libraries-Functions-Constants, echo = FALSE}
```



## Question 1


**1. Install the library "astsa" using the function: install.packages("astsa")**

```{r Question1-1, echo = -c(1:6)}
```

**2. Load the library: library(astsa)**

```{r Question1-2, echo = -c(1:6)}
```

**3. Use the function str() to see the information of a particular data series, such as
str(EQ5) for the Seismic Trace of Earthquake number 5 series.**

```{r Question3, echo = -c(1:6)}
```

**4. Plot the time series plots and histograms of the following 3 series. Feel free to use
the codes provided in the R scripts. Make sure that each of your graph has a title,
the axis ticks are clear, the axes are well-labelled, and use color intelligently.**

```{r Question1-4, echo = -c(1:6)}
```

**5. Write a few sentences to describe each of the series.**
*EQ5
*flue
*gas 
  
```{r Question1-5, echo = -c(1:6)}
```



**********

## Question 2

**Describe 3 examples you have used in your work or encounter in real life. Ideally, you
can even load at least one of these time series, plot it, and the write a few statements to
describe its characteristics.**

```{r Question2, echo = -c(1:9)}
```



**********

## Question 3

**Simulate a white noise series with 1000 random draws and plot (1) a time series plot and
(2) a histogram. The usual reqirements on graphics (described) in Question 1) applied.**


```{r Question3, echo = -c(1:4)}
```



**********

## Question 4

**Simulate (with 1000 random draws) two the following two zero-mean autoregressive model
with order 1 (i.e. AR(1)) models:**

$$yt=  0.9 y_{t1} + w$$
$$y_t=  0.2 y_{t1} + w$$

**Plot a time plot for each of the simulated series. Graph a histogram for each of thes
simulated series. Write a few statements to compare the two series.**

```{r Question4, echo = -c(1:2)}
```



**********

## Question 5

**Simulate (with 1000 random draws) the following 3 models:**

**1. A deterministic linear (time) trend of the form: yt = 10 + 0:5t**

```{r Question5-1, echo = -c(1:4)}
```

**2. Random walk without drift**

```{r Question5-2, echo = -c(1:4)}
```

**3. Random walk with drift = 0.5**

```{r Question5-3, echo = -c(1:4)}
```

**Plot a time plot for each of the simulated series. Graph a histogram for each of thes
simulated series. Write a few statements to compare the two series.**

```{r Question5-4, echo = -c(1:4)}
```

**********
